### 2015-02-18 doc/template/intro.tex

\chapter{Introduction}
\label{cha:intro}
The first contains a general introduction to the work. The goals are
defined and the modus operandi is explained.

\section{chap}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: