### 2015-05-19 doc/template/conclusion.tex
alles opslagen

\chapter{Conclusion}
\label{chapter5:conclusion}
The final chapter contains the overall conclusion. It also contains
suggestions for future work and industrial applications.

\section{trala}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: