### 2015-06-08 doc/template/optimum.tex
beetje overal verbeteren

\chapter{FlipIt game with virus propagation}
\label{chapter2:FlipIt with virus propagation}
%\documentclass[10pt]{article}
%\begin{document}

The optimum functions will be piecewise functions. There will be different optimum functions depending on $k_{D}$ and $d$. Two cases are considered.
\subsubsection{$k_{D} \leq d$}

\begin{description}
\item if $\delta_{A} < 2(k_{D} - d)$ \\
For case 2.A it is decreasing. Because $k_{D} \leq d$ the term $2(k_{D} - d)$ will always be negative so $\delta_{A}$ will also be negative. This means for case 1 and case 2.b that $\delta_{D} = 0$ and so the defender will not play.
\item $\delta_{A} = 2(k_{D} - d)$ \\
$\delta_{A} = 0$ so the attacker will not play. For case 1 and case 2.b the defender will also not play.
\item $\delta_{A} > 2(k_{D} - d)$ \\
For case 2.b it is decreasing. Case 2.a it is increasing. For case 1 $2(k_{D} - d)$ is always negative so the minimum is 0. This means that $\delta_{D} \in [0, \sqrt{d^{2} + 2\delta_{A}k_{D}}]$
\end{description}

\subsubsection{$k_{D} > d$}
\begin{description}
\item if $\delta_{A} < 2(k_{D} - d)$ \\
For case 2.A it is increasing. 
\item $\delta_{A} = 2(k_{D} - d)$ \\
\item $\delta_{A} > 2(k_{D} - d)$ \\
In every case the function is decreasing so the defender will not play.
\end{description}

From this analyses we can compute $opt_{D}(\delta_{A})$ for two different cases as:
 \begin{displaymath}
  opt_{D}(\delta_{A}) = \left\{
     \begin{array}{lr}
       0, & \delta_{A} < 2(k_{D} - d)\\
       0, & \delta_{A} = 2(k_{D} - d) \\
       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} > 2(k_{D} - d)
     \end{array}
   \right.
\end{displaymath}

For case :
 \begin{displaymath}
  opt_{D}(\delta_{A}) = \left\{
     \begin{array}{lr}
       0, & \delta_{A} < 2(k_{D} - d)\\
       0, & \delta_{A} = 2(k_{D} - d) \\
       \big[0,\sqrt{d^{2} + 2\delta_{A}k_{D}}\big], & \delta_{A} > 2(k_{D} - d)
     \end{array}
   \right.
\end{displaymath}
-------------------------------------------\\