### 2015-08-19 doc/template/nash.tex
even alles opslagen

\chapter{Optimal defence and attack strategies}
\label{chapter:Nash}
%\documentclass[10pt]{article}
%\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%			Introduction Chapter 1				%%%%%%
%%%%%												%%%%%%
%%%%%												%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%TODO-- rechtstreeks uit FlipIt paper --\\

In this chapter we are interested in finding the optimal defence and attack strategies. Ultimately, the optimal strategies can allow the determination of the Nash equilibria of the game. \\

%First the optimal functions are derived from the formulas in the previous chapter. From these piecewise functions we can derive the Nash equilibria. \\

Nash equilibria are points with the property that neither player benefits by deviating in isolation from the equilibrium. We can compute Nash equilibria for the periodic game as an intersection point of curves $opt_{D}$ and $opt_{A}$. 
\\
%As a second step, we are interested in finding Nash equilibria, points
%for which neither player will increase his benefit by changing his rate of play. 
More formally, a Nash equilibrium for the periodic game is a point $(\delta_{D}^{*},\delta_{A}^{*})$ such that
the defender's benefit $\beta_{D}(\delta_{D},\delta_{A}^{*}) $is maximised at $\delta_{D}= \delta_{D}^{*}$ and the attacker's benefit
$\beta_{A}(\delta_{D}^{*},\delta_{A}) $ is maximised at $\delta_{A}= \delta_{A}^{*}$.
To begin with, some useful notation. We denote by $opt_{D}(\delta_{A}$) the set of values (rates
of play $\delta_{D}$) that optimise the benefit of the defender for a fixed rate of play $\delta_{A}$ of the
attacker. Similarly, we denote by $opt_{D}(\delta_{D}$) the set of values (rates of play $\delta_{A}$) that optimise
the benefit of the attacker for a fixed rate of play $\delta_{D}$ of the defender. \\

% Fig \todo{fig} shows all the cases.\\
%\begin{figure}[hbtp]
%\centering
%\includegraphics[scale=0.4]{Images/bestresp.png}
%\caption{This figure shows the all the cases with their subcategories. 'A' stands for Case 2.A: $\delta_{D} \geq d+\delta_{A} \geq \delta_{A}$ and 'B' stands for Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $ }
%\label{grafiekbestr}
%\end{figure}

\section{Determining the piecewise functions $opt_{D}(\delta_{A})$}

To determine $opt_{D}(\delta_{A})$ we need to compute the derivative of  $\beta_{D}(\delta_{D},\delta_{A}) $ for a fixed $\delta_{A}$.
 We consider three cases for each piece of the piecewise function of $\beta_{D}$.

\subsection*{Case A: $\delta_{D} \leq d$}

The benefit formula obtained in the previous chapter is as follows:
\begin{equation}
\beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac{k_{D}}{\delta_{D}}
\end{equation}

The function is of the type $1-1/x$, see figure \ref{1x}. The root of the benefit function  and the root of the first derivative are as follows:
\begin{equation}
\beta_{D}(\delta_{D},\delta_{A}) =0  ~~~~~~ =>~~~~~~\delta_{D} = k_{D} \\
\end{equation}
\begin{equation}
\dfrac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ k_{D} = 0
\end{equation}

\begin{figure}[hbtp]
\centering
\includegraphics[scale=1]{Images/1x.png}
\caption{function of type $1-1/x$}
\label{1x}
\end{figure}

This means that if $\delta_{D} < k_{D}$  the function is negative and the defender will therefore not play, if $\delta_{D} > k_{D}$ the function is positive. If $k_{D} < 0$ the function will decrease and the defender will again not play (cost cannot be negative).  Assuming that costs are always non-negative $k_{D} > 0$, the function will increase, meaning that the slower the defender plays, the larger the benefit.

\subsection*{Case B: $ d \leq\delta_{D} \leq \delta_{A} + d$}
The defender's benefit formula obtained in the previous chapter in this case is as follows:
\begin{equation*}
\beta_{D}(\delta_{D},\delta_{A}) = 1 - \dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}} + \dfrac{d}{\delta_{A}}  - \dfrac{k_{D}}{\delta_{D}}
\end{equation*}

To know if the function decreases or increases we take the partial derivative of this formula for a fixed $\delta_{A}$:
\begin{equation*}\label{formdelta}
\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} = - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
\end{equation*}

The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
\begin{equation*}
\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
\end{equation*}

Given the sign of the coefficient of $\delta_{D}^{2}$, this
leads to the following deduction: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$. So we have a maximum at $\delta_{D} = min \{ \delta_{A} +d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ and $\delta_{D}$ has to be larger than $d$. Taking the minimum of the two values is needed because $\delta_{D}$ cannot be larger than $\delta_{A} +d$. \\
~~\\

\subsection*{Case C: $\delta_{D} \geq d+\delta_{A} $ }

The benefit formula for the defender obtained in the previous chapter in this case is as follows:
\begin{equation*}
\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} = \dfrac{\delta_{A} + 2 (d-k_{D})}{2\delta_{D}}
\end{equation*}

Given that $\delta_{D}$ is always positive, the benefit function can be either increasing or decreasing depending on the numerator of the above fraction. \\

For $\delta_{A} + 2(d-k_{D}) > 0$ the benefit will be always positive but decreasing, see figure \ref{ShapeUp}. 
The defender will always play as fast as he can if $\delta_{A} + 2(d-k_{D}) > 0$ for $k_{D} < d$ or $k_{D} > d$ because $\delta_{A}$ will be positive in either case. There is an edge case if $d=k_{D}$, which results in a benefit of $\beta_{D}(\delta_{D},\delta_{A})= \dfrac{\delta_{A}}{2\delta_{D}}$. \\
\begin{figure}
\centering
\includegraphics[scale=0.5]{Images/ShapesUp.png} 
\caption{The benefit function is of the shape of 1/x and is always decreasing if $\delta_{A} + 2(d-k_{D}) > 0$. }
\label{ShapeUp}
\end{figure}

For $\delta_{A} + 2(d-k_{D}) < 0$, the benefit will always be increasing but negative so the defender will not play. See figure \ref{ShapeDown}.  \\

\begin{figure}
\centering
\includegraphics[scale=0.5]{Images/ShapeDown.png} 
\caption{The benefit function is of the shape of -1/x and is always increasing if $\delta_{A} + 2(d-k_{D}) < 0$.}
\label{ShapeDown}
\end{figure}
%The derivative of the above formula for a fixed $\delta_{A}$ results in the following:
%\begin{equation*}
%\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} = -\dfrac{\delta_{A}}{2\delta_{D}^{2}} - \dfrac{d}{\delta_{D}^{2}} + \dfrac{k_{D}}{\delta_{D}^{2}}
%\end{equation*}
%The obtain the stationary points the first derivative is set equal to zero and the roots of the resulting equation are found:
%\begin{equation*}
%\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = 2(k_{D}-d) = dk_{D} - 2d
%\end{equation*}
%
%This leads to the following deduction:
%\begin{description}
%\item If $k_{D} \leq d$ 
%\begin{description}
% \item $\beta_{D}$ will be decreasing but always positive. If we minimize $\delta_{D}$ the value of $\beta_{D}$ will be higher. 
%\end{description}
%\item If $k_{D} > d$ 
%\begin{description}
%\item if $ \delta_{A} > 2(k_{D} -d)$, \\
%$\beta_{D}$ will be decreasing but always positive. If we minimize $\delta_{D}$ the value of $\beta_{D}$ will be higher. 
%\item if  $\delta_{A} = 2(k_{D} -d)$, \\
%the benefit of the defender will be $\beta_{D}=0$.
%\item if $\delta_{A} < 2(k_{D} -d)$, \\
%$\beta_{D}$ will be increasing but always negative. In this case the defender will not play. 
%\end{description}
%\end{description}
~~\\

%\subsection*{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $} 
%
%The benefit formula obtained in the previous chapter  (formula \ref{benfcase2b:defender}) for the defender in this case is as follows:
%\begin{equation*}
%\beta_{D}(\delta_{D},\delta_{A}) = \dfrac{\delta_{A}}{2\delta_{D}} + \dfrac{d}{\delta_{D}} - \dfrac{k_{D}}{\delta_{D}} - \dfrac{(d-(\delta_{D} - \delta_{A}))^{2}}{2\delta_{D}\delta_{A}}
%\end{equation*}
%
%The derivative of the above formula for a fixed $\delta_{A}$ results in the following:
%\begin{equation*}
%\beta_{D}(\delta_{D},\delta_{A}) =  - \dfrac{1}{2\delta_{A}} + \dfrac{k_{D}}{\delta_{D}^{2}} + \dfrac{d^{2}}{2\delta_{D}^{2}\delta_{A}}
%\end{equation*}
%
%
%The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
%
%\begin{equation*}
%\frac{\partial \beta_{D}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D} = \sqrt{2\delta_{A}k_{D} + d^{2}}
%\end{equation*}
%
%
%For case 2.B this leads to the following deduction, which results in the same formula as for case 1 but with a small difference for the value of $\delta_{D}$: The function increases on $[0, \sqrt{2\delta_{A}k_{D} + d^{2}}]$ and is decreasing on $[\sqrt{2\delta_{A}k_{D} + d^{2}}, \infty]$. Because $\delta_{D} \geq \delta_{A}$ there is a maximum on $\delta_{D} = maximum \{ \delta_{A}, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $ or because $d+\delta_{A} \geq \delta_{D}$ there is also on $\delta_{D} = minimum \{ \delta_{A}+d, \sqrt{2\delta_{A}k_{D} + d^{2}} \} $. %-Remark- \todo{beter uitschrijven, deltaD moet tussen die twee waarden liggen}\\

\section{Best responses of $\delta_{D}$}
The optimum functions will be piecewise functions. We distinguish six cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
%We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates. 
Figure XXX shows on the left how the three cases can be combined in a single continuous function (shown as the thick line). On the right, the derivative function indicates at which points the benefit will change. It is interesting to note that the derivative is also continuous, meaning that the benefit function is smooth and has no rough edges or corners. This implies that there are no sudden changes in benefit for small changes of period.

\subsection*{$\delta_{A} < 2(k_{D} - d)$}
For case A, the function will always increase if $k_{D} > 0$. If $k_{D} < 0$ it means that the cost is negative and we assume a positive cost. The function will thus never decrease. $k_{D} =0$ is an edge case and will be covered later in the paragraph of edge cases. So for case A we can conclude that the function is always increasing. In the next case, case B, the benefit function will increase in the interval $[0, 2(k_{D}-d)[$. In case C, for $\delta_{A} < 2(k_{D} - d)$, the benefit function will always increase. Because for all the cases we are dealing with a negative benefit function that is increasing, the maximum benefit is achieved for $\delta_{} = \infty$. This can be seen in (A) in figure \ref{case1D}.

\begin{figure}[hbtp]
\centering
\includegraphics[scale=1]{Images/case1b.pdf}
\caption{ Illustration of the best response functions. The left functions are the piecewise benefit function of the defender for a certain values of $k_{D},d$ and $\delta_{A}$.. The right functions are the derivatives of the benefit functions on the left. The best responses are the values where the function intersects with the x-as. (A): $\delta_{A} < 2(k_{D}-d)$, (B): $\delta_{A} = 2(k_{D} - d)$, (C): $\delta_{A} > 2(k_{D} - d)$}
\label{case1D}
\end{figure}

%From case C above, it follows that if $\delta_{A} < 2(k_{D} - d)$ the benefit is increasing but it will always be non-positive. The defender will try not to play ($\delta_{D}= \infty$).

%From case C it follows that from interval $[2(k_{D} -d), \sqrt{2k_{D}\delta_{A} + d^{2}}]$  (because for case 2 $\delta_{D} \geq \delta_{A}$), that the benefit is increasing.  The maximum of case 2.a and the maximum of case 2.b can be brought together. It follows that the optimal benefit is achieved at the $\delta_{D}=minimum[\delta_{A} + d, \sqrt{2k_{D}\delta_{A} + d^{2}}]$, which is $\delta_{D}=\sqrt{2k_{D}\delta_{A} + d^{2}}$. But because of case 2.a, the defender's maximum benefit is not to play at all.

\subsection*{$\delta_{A} = 2(k_{D} - d)$}
The delay can not be a negative value. So for $\delta_{A}$, a positive period, to be equal to  $2(k_{D} - d)$, $k_{D}$ has to be positive. \\
For case A, if $k_{D}>0$ the benefit function is always increasing. In case B, the value of $\delta_{A} =2(k_{D} - d)$ has to filled in in the root, $\sqrt{2k_{D}\delta_{A} + d^{2}}$. This gives the value $2k_{D}-d$. $2(k_{D}-d) < 2k_{D}-d$ so this means that $\delta_{A}$ is smaller than $ \sqrt{2k_{D}\delta_{A} + d^{2}}$, so the function is increasing. A maximum is achieved for $\delta_{D}$ equal to the maximum of $\left\lbrace \delta_{A}+d, \sqrt{2k_{D}\delta_{A} + d^{2}}\right\rbrace $. If we fill in $\delta_{A} = 2(k_{D} - d)$, the two values are the same so they are both a maximum.
If we fill in the value of $\delta_{A} = 2(k_{D} - d)$ in case C, the benefit will always be 0. Case C is the case for all values $\delta_{D} \geq \delta_{A} + d$, so $\beta_{D}=0$ is true for all $\delta_{D} \in [\delta_{A}+d, \infty]$.
The defender's maximum benefit is thus achieved for all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$. See (B) in figure \ref{case1D}.

%For case C, $\beta_{D}(\delta_{D}, \delta_{A})=0$ for  all $\delta_{D} \in [\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$. For case B, the benefit is increasing in the interval $[\delta_{A},\sqrt{2k_{D}\delta_{A} + d^{2}}]$. If we put both cases back together, we get a maximum at $\sqrt{2k_{D}\delta_{A} + d^{2}}$. So in this case the defender's maximum benefit is achieved for any $\delta_{D}$ in $[\sqrt{2k_{D}\delta_{A} + d^{2}}, \infty]$ with value 0.

\subsection*{$\delta_{A} > 2(k_{D} - d)$ }
Again for case A, we assume that $k_{D}$ is positive which implies that the benefit function for the defender is always increasing. If $\delta_{A} > 2(k_{D} - d)$, the benefit function of the defender in case B will only increase in the interval $]2(k_{D} - d), \sqrt{2k_{D}\delta_{A} + d^{2}}$. The maximum is achieved for $\delta_{D}$ equal to the maximum of $\left\lbrace \delta_{A}+d, \sqrt{2k_{D}\delta_{A} + d^{2}}\right\rbrace $. If we fill in $\delta_{A} > 2(k_{D} - d)$ , $\sqrt{2k_{D}\delta_{A} + d^{2}}$ is the maximum value. In case C, for all $\delta_{A} > 2(k_{D} - d)$, the benefit function is always decreasing. If we put all the pieces together, the defender's maximum benefit is achieved for $\delta_{D} = \sqrt{2k_{D}\delta_{A} + d^{2}}$. See (C) in figure \ref{case1D}.

%For case C, it follows that the benefit is decreasing but positive. So the defender will try to play as fast as possible ($\delta_{D} \rightarrow 0$). For case B it is increasing in the interval $[ 0,\sqrt{2k_{D}\delta_{A} + d^{2}}]$. It follows that the defender's maximum benefit is achieved for $\delta_{D} = \sqrt{2k_{D}\delta_{A} + d^{2}} $  \\

~~\\
From this analyses we can compute $opt_{D}(\delta_{A})$ : \\

 \begin{displaymath}
  opt_{D}(\delta_{A}) = \left\{
     \begin{array}{lr}
          \infty , & \delta_{A} < 2(k_{D} - d)\\
      \left[ \sqrt{2k_{D}\delta_{A} + d^{2}},\infty\right] , & \delta_{A} = 2(k_{D} - d) \\
      \sqrt{2k_{D}\delta_{A} + d^{2}}, & \delta_{A} > 2(k_{D} - d)
     \end{array}
   \right.
\end{displaymath}

~~\\

\subsection*{Edge cases:}

\subsection*{$\delta_{A}=0$ and $k_{D}=0$}

If $k_{D}=0$ it means that every flip of the defender is free. So for the cost it does not matter for the defender if he plays faster or slower. $\delta_{A}=0$ means that the attacker will play as fast as he can.  
\begin{itemize}
\item if $d=0$, if we look at $opt_{D}(\delta_{A})$ it means that we are in the second piece. The defender will play with a rate of $[0,\infty]$. 
\item if $d >0$, we have to look at the third piece of  $opt_{D}(\delta_{A})$ . Here it follows that the defender has to play with a rate of \textit{d}.
\end{itemize}

Combining those two leads to a rate for the defender equal to $[0,d]$. This means that the defender will never not play, which is intuitively correct if we know that the defender has no disadvantage to play because there is no cost involved.\\

Or more intuitively: If $\delta_{A}=0$ it means that the attacker is playing as fast as possible. For the defender it does not cost anything to flip, because $k_{D}=0$. We know that if $\delta_{D} \leq d$ the defender always has a benefit of 1. This means that for this case the defender will play with a rate $\delta_{D} \in [0,d]$.\\
\subsection*{$\delta_{A}=0$}
If $\delta_{A}=0$ it means that the attacker is playing as fast as possible. The defender has still a cost for every flip. This means we have to look at different values for $k_{D}$ and $d$.
\begin{itemize}
\item if $k_{D} > d$, it follows that the defender will not play (rate equal to $\infty$). This is already the case for the first piece in the piecewise function. $\delta_{A}$ will always be smaller than $2(k_{D}-d)$ if $k_{D} > d.$
\item if $k_{D}=d$, it corresponds to the second piece: the defender will play with a rate of $[d,\infty]$.
\item if $k_{D} < 0$, this corresponds with the last piece, where the rate of the defender is equal to \textit{d}.
\end{itemize}
By combining the last two pieces we have  $d ~~$for$~~ \delta_{A} =0 ~~$ and$ ~~k_{D} \leq d$.\\

\subsection*{$k_{D}=0$}
The cost of flipping for the defender is equal to zero. If we look at the benefit function of the defender and it's derivative in figure \ref{cost0}, we can see that the benefit for case B increases until $d$ and that the benefit in case A stays the same until $d$. After point $d$ the benefit decreases in all the cases.

\begin{figure}[hbtp]
\centering
\includegraphics[scale=1]{Images/cost0.pdf} 
\caption{Benefit function and derivative for a cost $k_{D} = 0$ and with $d (=2)$ and $\delta_{A}=(1.5)$ not equal to zero.}
\label{cost0}
\end{figure}

\subsection{Conclusion}
Including the edge cases, $opt_{D}(\delta_{A})$ is as follows:  \\

 \begin{displaymath}
  opt_{D}(\delta_{A}) = \left\{
     \begin{array}{lr}
     \left[0,d\right] & \delta_{A} =0 ~~\& ~~k_{D}=0 \\
     \left[0,d\right] & k_{D}=0\\
     d & \delta_{A} =0 ~~ \& k_{D} \leq d\\
          \infty , & \delta_{A} < 2(k_{D} - d)\\
      \left[ \sqrt{2k_{D}\delta_{A} + d^{2}},\infty\right[ , & \delta_{A} = 2(k_{D} - d) \\
      \sqrt{2k_{D}\delta_{A} + d^{2}}, & \delta_{A} > 2(k_{D} - d)
     \end{array}
   \right.
\end{displaymath}
%*****************************************************************
%
% optimum functies voor beta A
%
%*********************************************************************
\section{Determining the piecewise functions $opt_{A}(\delta_{D})$}
%To start with we only consider the case where $d < \delta_{D}$, because if $d > \delta_{D}$ the benefit of the defender \todo{nakijken, def of att} is always 1. \\
To determine the optimal strategy of the attacker we also need to determine $opt_{A}(\delta_{D})$ by computing the derivative of $\beta_{A}(\delta_{D},\delta_{A})$ for a fixed $\delta_{D}$. We consider the three cases of the piecewise function of $\beta_{A}$: \\

\subsection*{Case A: $\delta_{D} \leq d$}

The benefit formula for the attacker for this case obtained in the previous chapter is as follows:
\begin{equation}
\beta_{A}(\delta_{D},\delta_{A}) =  - \dfrac{k_{A}}{\delta_{A}}
\end{equation}

The function is of the type $-1/x$, see figure \ref{1xx}. The root of the benefit function is as follows:
\begin{equation}
\beta_{A}(\delta_{D},\delta_{A}) =0  ~~~~~~ =>~~~~~~k_{D} = 0 \\
\end{equation}

\begin{figure}[hbtp]
\centering
\includegraphics[scale=1]{Images/1x.png}
\caption{function of type $-1/x$}
\label{1xx}
\end{figure}

This means that the optimal benefit is achieved for a value $\delta_{A}=0$.

\subsection*{Case B: $ \delta_{A} \leq \delta_{D} -d$ }
The benefit formula for the attacker for this case obtained from the previous chapter is as follows:
\begin{equation*}
\beta_{A}(\delta_{D},\delta_{A}) =1- \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{k_{A}}{\delta_{A}} - \dfrac{d}{\delta_{D}}
\end{equation*}
The derivative for a fixed $\delta_{D}$ is as follows:
\begin{equation*}
\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = \dfrac{-1}{2\delta_{D}} + \dfrac{k_{A}}{\delta_{A}^{2}}
\end{equation*}
The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
\begin{equation*}
\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{A} = \sqrt{2\delta_{D}k_{A}}
\end{equation*}
It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing on $[0,\sqrt{2k_{A}\delta_{D}}]$ and decreasing on $[\sqrt{2k_{A}\delta_{D}}, \infty]$ and thus has a maximum on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $. The minimum between $\delta_{D}-d$ and $ \sqrt{2k_{A}\delta_{D}}$ is needed because $\delta_{A} $ cannot exceed $\delta_{D}-d$ in this case. \\

%\subsection*{Case 2.B: $d+\delta_{A} \geq \delta_{D} \geq  \delta_{A} $} 
%
%The benefit formula obtained from the previous chapter \ref{benfcase2b:attacker} for this case is as follows: 
%\begin{equation*}
%\beta_{A}(\delta_{D},\delta_{A}) = 1 - \delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{(d-(\delta_{D}-\delta_{A})^{2}}{2\delta_{D}\delta_{A}} 
%\end{equation*}
%The derivative for a fixed $\delta_{D}$ is as follows:
%\begin{equation*}
%\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}}
%\end{equation*}
%The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
%\begin{equation*}
%\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~ \delta_{D}= \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
%\end{equation*}
%
%
%It follows that $\beta_{A}(\delta_{D},\cdot)$ is increasing and non-positif if $\delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$ and decreasing and positive if $\delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}$. This is the same result as in case 1.\\

\subsection*{Case C: $d \leq \delta_{D} -d \leq \delta_{A} $}

The benefit formula for the attacker for this case obtained in the previous chapter is as follows:
\begin{equation*}
\beta_{A}(\delta_{D},\delta_{A}) =\dfrac{\delta_{D}}{2\delta_{A}} - \dfrac{k_{A}}{\delta_{A}} + \dfrac{d^{2}}{2\delta_{D}\delta_{A}} - \dfrac{d}{\delta_{A}}
\end{equation*}
The derivative for a fixed $\delta_{D}$ is as follows:
\begin{equation*}
\dfrac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{A}} = -\dfrac{\delta_{D}}{2\delta_{A}^{2}} + \dfrac{k_{A}}{\delta_{A}^{2}} - \dfrac{d^{2}}{2\delta_{D}\delta_{A}^{2}} + \dfrac{d}{\delta_{A}^{2}} = \dfrac{-\delta_{D}^{2} - d^{2} + 2\delta_{D}d + 2\delta_{D}k_{A}}{2\delta_{A}^{2}\delta_{D}}
\end{equation*}
The stationary points (maximum, minimum) can be found by setting the first derivative equal to zero and finding the roots of the resulting equation:
\begin{equation*}
\frac{\partial \beta_{A}(\delta_{D},\delta_{A})}{\partial \delta_{D}} =0 ~~~~~~ =>~~~~~~  \delta_{D}^{2}-\delta_{D}(2k_{A}-2d) + d^{2}= 0
\end{equation*}

The roots of this quadratic equation are $\delta_{D}=d+k_{A}\pm \sqrt{2dk_{A}+k_{A}^{2}}$. Because for this case $d\leq \delta_{D}$, we only look at root $\delta_{D}=d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$. $\sqrt{2dk_{A}+k_{A}^{2}}$ is bigger than $k_{A}$ so if we subtract it from $d+k_{A}$, we are in the part where $delta_{D}$ is smaller than d. \\
It follows that $\beta_{A}(\delta_{D},\delta_{A})$ is increasing and non-positive if $\delta_{D}> d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$ and decreasing and positive if $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$  \\

\section{Best responses of $\delta_{A}$}
The optimum function of the attacker is found in the same way as the optimum function of the defender. The derivatives of each piece of the piecewise benefit function of the attacker is taken. From these derivatives the best responses are derived. \\
Case A is an edge case and will be addressed in the paragraph about the edge cases.
%The optimum functions will be piecewise functions. We distinguish three cases for different values of $\delta_{D}$ regarding $d$ and $k_{A}$. 
%
%
%For this term $\dfrac{(\delta_{D}-d)^{2}}{2k_{A}} $ , $d$ has to be bigger than  $\delta_{D}$ because the cost $\delta_{D}$ cannot be negative. This was an assumption that was already made, because the benefit of the defender will always be 1 if $d$ is bigger than  $\delta_{D}$.

\subsubsection*{$\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$} 
For case B, the function is increasing. The function has a maximum on on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $. For $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the minimum of the two values is $\delta_{D}-d$.\\
For case C, if $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the benefit function is increasing. Because we have a continuous function and the benefit is still increasing and negative in case C, the defender will try not to play. The defender's optimum benefit is achieved in $\delta_{A}=\infty$. See (A) in figure \ref{case1A}.

\begin{figure}[hbtp]
\centering
\includegraphics[scale=1]{Images/case1A.pdf}
\caption{ Illustration of the best response functions. The left functions are the piecewise benefit function of the attacker for a certain values of $k_{A},d$ and $\delta_{D}$. The right functions are the derivatives of the benefit functions on the left. The best responses are the values where the function intersects with the x-as. (A): $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (B): $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, (C): $\delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$}
\label{case1A}
\end{figure}

%From case 1 and case 2.b above, it follows that the benefit is decreasing but positive, which means that the attacker will play as fast as he can. For case 2.a  $\delta_{D} \geq \delta_{A}$ is increasing. It follows that for this case the attacker's maximum benefit is achieved for $\delta_{A} = \sqrt{d k_{A}\delta_{D}}$.
%. In this case the attacker's maximum benefit is reached at $\delta_{A} = \delta_{A} = \sqrt{dk_{A}\delta_{D}}$  .
\subsubsection*{$\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$} 
For case B, the function is increasing. The function has a maximum on on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $. For $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the minimum of the two values is $\delta_{D}-d$.\\
For case C, $\delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$ is a root so the benefit function of the attacker will be equal to zero. Because case C is the case where $\delta_{A}$ has to be bigger or equal to $\delta_{D}-d$, $\beta_{A}=0$ is valid for all $\delta_{A} \in [\delta_{D}-d,\infty]$.\\
Putting the pieces together gives us the optimum benefit for the attacker for all $\delta_{A} \in [\delta_{D}-d,\infty]$. See (B) in figure \ref{case1A}.

%The benefit $\beta_{A}(\delta_{D},\delta_{A})=0$ for case 1 and case 2.b for all $\delta_{D}$ in $[0, \infty [$. For case 2.a it follows that the benefit is increasing. The attacker's maximum benefit is achieved at any $\delta_{A} $ in 
\subsubsection*{$\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$} 
For case B, the function is increasing. The function has a maximum on on $\delta_{A} = minimum \{\delta_{D} -d, \sqrt{2k_{A}\delta_{D}} \} $. For $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$, the minimum of the two values is $\delta_{D}-d$.\\
For case C, if $\delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}$ the function is decreasing. So the optimum benefit for all the cases is when $\delta_{D} = \delta_{D}-d$. See (C) in figure \ref{case1A}.\\
%For case 1 and case 2.b the benefit is increasing but non-positive. This means that the attacker won't play ($\delta_{A} = \infty$).

From this analyses we can compute $opt_{A}(\delta_{D})$ :

 \begin{displaymath}
  opt_{A}(\delta_{D}) = \left\{
     \begin{array}{lr}
            \infty & \delta_{D} < d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}} \\
       \left[ \delta_{D}-d, \infty\right],  & \delta_{D} = d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}} \\
    	\delta_{D}-d, & \delta_{D} > d+k_{A} + \sqrt{2dk_{A}+k_{A}^{2}}
     \end{array}
   \right.
\end{displaymath}
\\

\subsection*{Edge cases:}

\subsection*{$\delta_{D}=0$ and $k_{A}=0$}
If $\delta_{D}=0$, looking at the benefit formula of the attacker, it follows that the period of the defender will always be equal or smaller than the delay. The delay cannot be negative. So from case $\delta_{D} \leq d$ it follows that the attacker will always have a benefit equal to zero, because $k_{A}=0$. For the attacker it does not matter what he plays. The benefit will always be the same. The optimal strategy in this case for the attacker will be playing at a rate $\delta_{A} \in [0,\infty ]$.

\subsection*{$k_{A}=0$}
If the cost of playing for the attacker is zero, the attacker can play as fast as he can.  If $\delta_{D} \leq d$, looking at the benefit formula for the attacker, it does not matter what the attacker does, his benefit will always be zero. He will play with a rate $\delta_{A} \in [0,\infty ]$.  We can merge this with the previous case where $\delta_{D}=0$ and $k_{A}=0$.  If $\delta_{D} > d$, the attacker will play. He will play with a rate of $\delta_{A} \in [0, \delta_{D} - d ]$. 
\subsection*{$\delta_{D}=0$}
If $\delta_{D}=0$ it follows from the case $\delta_{D} \leq d$ that the attacker will always have a negative benefit, unless the cost is zero. If the cost is zero we have again the case of $\delta_{D}=0$ and $k_{A}=0$. But if the cost is non-zero, the optimal strategy of the defender is not moving at all.

\subsection{Conclusion}
Including the edge cases yields the following for $opt_{A}(\delta_{D})$ : \\

 \begin{displaymath}
  opt_{A}(\delta_{D}) = \left\{
     \begin{array}{lr}
     \infty & \delta_{D} = 0 \\
     \left[0,\infty\right] & \delta_{D} \leq d ~~\& ~~k_{A}=0 \\
          \left[0,\delta_{D} - d\right] & \delta_{D} > d ~~\& ~~k_{A}=0 \\
                 \infty & \delta_{D} < \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
       \left[ \sqrt{d k_{A}\delta_{D}}, \infty\right],  & \delta_{D} = \dfrac{(\delta_{D}-d)^{2}}{2k_{A}} \\
       \sqrt{d k_{A}\delta_{D}}, & \delta_{D} > \dfrac{(\delta_{D}-d)^{2}}{2k_{A}}
     \end{array}
   \right.
\end{displaymath}
\\

\section{Conclusion}

The piecewise functions $opt_{D}$ and $opt_{A}$ give some interesting results. \todo{resultaten nog toevoegen, zelfde als in de andere conclusie}

The Nash Equilibria can be found as the intersection points of the piecewise functions $opt_{A}(\delta_{D})$ and $opt_{D}(\delta_{A})$. For this we have to compare the function in terms of the relationship of the players move cost. We distinguish three cases: $k_{A} < k_{D} , ~k_{A} > k_{D} $ and $k_{A} = k_{D}$.